The proof equivalence problem for multiplicative linear logic is \textsc{pspace}-complete v0.3

### 2014-01-12 main/s_level_0.tex
added octopus roll lemma

@@

\begin{proof}
For a sequent with 1 tensor-formula, the problem is reduced to syntactic equality.
%
For a sequent with 2 or more tensor-formulae, by Propositions~\ref{prop:level0 max binary} and~\ref{prop:level0 min binary} the equivalence of two nets is determined by their relative parity.
%
Following Definition~\ref{def:combinatorial permutation} the parity of a net can be read off in a single traversal of the net.
%
This yields a linear-time algorithm.
%
\end{proof}

>>>>

\begin{proposition}
\label{prop:parity determines equivalence}
Two proof nets for a 1-alternation sequent with at least two tensor-formulae are equivalent if and only if they have the same parity.
\end{proposition}

@@

\begin{proof}
For a sequent with 1 tensor-formula, the problem is reduced to syntactic equality.
%
For a sequent with 2 or more tensor-formulae, by Propositions~\ref{prop:parity determines equivalence} the equivalence of two nets is determined by their parity.
%
Following Definition~\ref{def:combinatorial permutation} the parity of a net can be read off in a single traversal of the net.
%
This yields a linear-time algorithm.
%
\end{proof}

\begin{lemma}
\label{lem:double exchange}
Let $\links$ be a proof net where $\links(a)=(u)$, $\links(b)=v$, $\links(c)=x$, and $\links(d)=y$, and for any switching the path from $u$ to $y$ passes through the links $b-v$ and $x-c$.
%
Then $\links\sim\links'$ where $\links'(a)=v$, $\links'(b)=u$, $\links'(c)=y$, and $\links'(d)=x$, and $\links'(z)=\links(z)$ otherwise.
\end{lemma}

\begin{proof}
\[
	\vc{\doubleExchange uvxy}
	\sim
	\vc{\doubleExchange uyxy}
	\sim
	\vc{\doubleExchange uyxu}
	\sim
	\vc{\doubleExchange vyxu}
	\sim
	\vc{\doubleExchange vyyu}
	\sim
	\vc{\doubleExchange vuyu}
	\sim
	\vc{\doubleExchange vuyx}
\]
\end{proof}

### 2014-01-12 main/s_level_2.tex
added octopus roll lemma

The interpretation $\itn\gamma$ of a total configuration $\gamma$ for a graph $G$ is a linking $\links$ constructed incrementally, for each successive edge $e$, and for each successive weight element $W$ within $e$, as follows.
%
Let the rightmost $1$-occurrence of each vertex-gadget $\itn v$ be named $\alt v$, and the rightmost $\bot$-occurrence of each edge-gadget $\itn e$ named $\alt e$:
\[	
	\itn v = C\parr\dotso\parr C\parr \named{\alt v}1 \qquad \itn e= W\tn\dotso\tn W\tn\named{\alt e}\bot~.
\]
To define $\itn\gamma$, firstly let $\itn\gamma(e)=\alt v$ where $\gamma(e)=v$.
%
Then successively for each edge $e$ and each weight element $W$ in $e$, if $\gamma(e)$ has a first free constraint element $C$, extend $\itn\gamma$ to include the natural linking on $W,C$; otherwise, extend $\itn\gamma$ by the natural linking on the sequent consisting of $W$ plus the first $3n+4$ free edge absorbers.

@@

>>>>

The rightmost $1$-occurrence of each vertex-gadget $\itn v$, named $\alt v$ below, and the rightmost $\bot$-occurrence of each edge-gadget $\itn e$, named $\alt e$, will be called the \emph{indicator vertices} of $\itn v$ and $\itn e$.
\[	
	\itn v = C\parr\dotso\parr C\parr \named{\alt v}1 \qquad \itn e= W\tn\dotso\tn W\tn\named{\alt e}\bot~.
\]

\begin{definition}
\label{def:configuration interpretation}
The \emph{interpretation} $\itn\gamma$ of a total configuration $\gamma$ for a graph $G$ is a linking $\links$ constructed incrementally, for each successive edge $e$, and for each successive weight element $W$ within $e$, as follows.
%
Let $\gamma(e)=v$; firstly, the the indicator vertex of $\itn e$ links to the indicator of $\itn v$.
%
Then successively for each weight element $W$ in $e$, if $\itn v$ has a first free constraint element $C$, extend $\itn\gamma$ to include the natural linking on $W,C$; otherwise, extend $\itn\gamma$ by the natural linking on the sequent consisting of $W$ plus the first $3n+4$ free edge absorbers.
\end{definition}

@@

In a proof net for $\itn G$, an edge-gadget $\itn e$ that is in the empire of an appropriate vertex-gadget $\itn v$ is \emph{naturally linked} if the indicator of $\itn e$ connects to the indicator of $\itn v$, and each weight element $W$ of $\itn e$ is either naturally linked to a constraint element $C$ in $\itn v$ or linked only to edge-absorbers.

\begin{lemma}
\label{lem:octopus roll}
In a proof net $\links$ for a sequent $\Gamma=\itn v,\itn{e_1},\dotsc,\itn{e_m},\bot^p$ where each edge-gadget is naturally linked, if a weight element $W_i$ in $\itn{e_i}$ is linked to $C$ in $\itn v$ and $W_j$ in $\itn{e_j}$ is linked to edge-absorbers $\bot^n$, then there is a net $\links'\sim\links$ in which $W_j$ is naturally linked to $C$, $W_i$ is linked to $\bot^n$, and $\links'$ agrees with $\links$ otherwise.
\end{lemma}

\begin{proof}
Let $W_i=A\parr B\parr C$, $W_j=D\parr E\parr F$, and $C=X\tn Y$.
%
We will illustrate the path of permutations for the case where $A,B,X$ and $D,X$, and thus also $C,Y$ and $E,F,Y$, are balanced sequents; other cases are similar.
%
\begin{enumerate}
	\item
The initial configuration is illustrated below; other weight and constraint elements are omitted, and $v$, $i$, and $j$ are the indicator vertices of $\itn v$, $\itn{e_i}$, and $\itn{e_j}$ respectively.
\[
	\octorollA1
\]
	\item
The link $i-v$ is re-attached to connect to an edge-absorber together with only the links from the first $\bot$ of each of $D$, $E$, and $F$. 
\[
	\octorollB2
\]
	\item
Secondly, the link from the first $\bot$ of $D$ is moved to $X$, and those of $E$ and $F$ are moved to $Y$.
\[
	\octorollB3
\]
	\item\label{item:exchange added}
There are subnets over the sub-sequents $A,B,X,D,\bot^m$ and $C,Y,E,F,\bot^k$.
%
These subnets may be rewired so that $D\parr E\parr F$ is naturally linked to $X\tn Y$: by Proposition~\ref{prop:parity determines equivalence}, the resulting subnets are equivalent as long as their parity is preserved.
%
Two links from $C$ to $X$ should remain exchanged, compared to the natural linking, for step \ref{item:exchange needed} below.
%
The links of $A,B,C$ connect to the edge-absorbers $\bot^m$ and $\bot^k$, with one remaining link from $B$ to $X$ and one from $C$ to $Y$.
\[
	\octorollC
\]
	\item
The link from $C$ to $Y$ is moved towards an edge-absorber connected to $A,B$.
\[
	\octorollD1
\]
	\item\label{item:exchange needed}
The link from $B$ to $X$ is the one remaining connection between the edge-gadgets $\itn{e_i}$ and $\itn{e_j}$.
%
Lemma~\ref{lem:double exchange} allows to swap the targets of the link from $B$ to $X$ and the link from $i$, and simultaneously undo the exchange in the links from $C$ to $X$ added in step \ref{item:exchange added} above.
\[
	\octorollD2
\]
	\item
The link from $i$ may be re-attached to $v$ to yield the final configuration.
\[
	\octorollA2
\]
\end{enumerate}
\end{proof}
